### 2017-02-27 readme.Rmd
update readme

There are two main functions in the `qrbp` package, the first and main function can be used to generate `n` quasi-random background points within a spatial domain. The for background points can be used in
Poisson point process modelling in R. The main function is \code{qrbp}, which takes some coordinates, an area of interest and covariates (and other parameters) and produces a \code{qrbp} dataset. It is essentially a wrapper around the (excellent and existing) \code{\link[MBHdesign]{quasiSamp}} function in the MBHdesign package which already implements quasi-random sampling within a basic domain.

@@

For a laugh, let's generate some quasirandom background points and plot them against the presence points
```{r}
library(qrbp)
POdata <- species[species$Occurrence == 1,]
bkpts <- generate_background_points(number_of_background_points = 1000,
                                    known_sites = POdata@coords,
                                    study_area = preds[[1]],
                                    model_covariates = preds,
                                    method = 'quasirandom')

Now let's plot our background points
```{r}
plot(preds[[1]])
points(bkpts[bkpts$presence == 0,c("x","y")],col='blue',pch=16,cex=.5)
points(bkpts[bkpts$presence == 1,c("x","y")],col='red',pch=16,cex=1)
```

Now let's try and generate a ppm using a poisson gam
```{r, warning=FALSE}
library(mgcv)
fm1 <- gam(presence ~ s(elevation) +
              s(precipitation) +
              s(temperature) +
              s(vegetation) + 
              offset(log(weights)),
              data = bkpts,
              family = poisson())

p <- predict(object=preds,
             model=fm1,
             type = 'response',
             const=data.frame(weights = 1))

p_cell <- p*res(preds)[1]*res(preds)[2]

plot(p_cell)
```

Let's compare to a PA sdm - hopefully we are in the right ball park
```{r,message=FALSE,error=FALSE,warning=FALSE}
d <- sdmData(formula= ~., train=species, predictors=preds)
m1 <- sdm(Occurrence~.,data=d,methods=c('gam'))
p1 <- predict(m1,newdata=preds)
plot(p1)
```
### References 
Grafström, Anton, and Yves Tillé. "Doubly balanced spatial sampling with spreading and restitution of auxiliary totals." Environmetrics 24.2 (2013): 120-131.

>>>>

There are two main functions in the `qrbp` package, the first and main function can be used to generate background points within a spatial domain. Generation of background points can be used in Poisson point process modelling in R. The main function is \code{generate_background_points}, which takes a Raster* as a study area, covariates - which are a Raster stack on the of the same resolution and extent as the study area. It is essentially a wrapper around the (excellent and existing) \code{\link[MBHdesign]{quasiSamp}} function in the MBHdesign package which already implements quasi-random sampling within a basic domain.

For a laugh, let's generate some quasirandom background points and plot them against the presence points
```{r}
library(qrbp)
POdata <- species[species$Occurrence == 1,]
bkpts_quasi <- generate_background_points(number_of_background_points = 10000,
                                    known_sites = POdata@coords,
                                    study_area = preds[[1]],
                                    model_covariates = preds,
                                    resolution = 2000,
                                    method = 'quasirandom_covariates')
```

For a laugh, let's generate some quasirandom background points and plot them against the presence points
```{r}
library(qrbp)
POdata <- species[species$Occurrence == 1,]
bkpts_grid <- generate_background_points(number_of_background_points = 10000,
                                    known_sites = POdata@coords,
                                    study_area = preds[[1]],
                                    model_covariates = preds,
                                    resolution = 2000,
                                    method = 'grid')
nrow(bkpts_grid)
```

Now let's plot our background points
```{r}
par(mfrow=c(1,2))
plot(preds[[1]])
points(bkpts_quasi[bkpts_quasi$presence == 0,c("x","y")],col='blue',pch=16,cex=.5)
points(bkpts_quasi[bkpts_quasi$presence == 1,c("x","y")],col='red',pch=16,cex=1)
plot(preds[[1]])
points(bkpts_grid[bkpts_grid$presence == 0,c("x","y")],col='pink',pch=16,cex=.2)
points(bkpts_grid[bkpts_quasi$presence == 1,c("x","y")],col='red',pch=16,cex=1)

Now let's try and generate a ppm using a poisson gam
```{r, warning=FALSE}
library(mgcv)
fm1 <- gam(presence ~ s(elevation) +
              s(precipitation) +
              s(temperature) +
              s(vegetation) + 
              offset(log(weights)),
              data = bkpts_quasi,
              family = poisson())

p1 <- predict(object=preds,
             model=fm1,
             type = 'response',
             const=data.frame(weights = 1))

p1_cell <- p1*(res(preds)[1]*res(preds)[2])

par(mfrow=c(1,2))

plot(p1_cell)

fm2 <- gam(presence ~ s(elevation) +
              s(precipitation) +
              s(temperature) +
              s(vegetation) + 
              offset(log(weights)),
              data = bkpts_grid,
              family = poisson())

p2 <- predict(object=preds,
             model=fm2,
             type = 'response',
             const=data.frame(weights = 1))

p2_cell <- p2*(res(preds)[1]*res(preds)[2])

plot(p2_cell)

Let's compare the spatial prediction of the PPM against the PA species distribution model - hopefully we are in the right ball park
```{r,message=FALSE,error=FALSE,warning=FALSE}
d <- sdmData(formula= ~., train=species, predictors=preds)
dat <- cbind(species$Occurrence,d@features)
colnames(dat)[1]<-'occurrence'
fm3 <- gam(occurrence ~ s(elevation) +
              s(precipitation) +
              s(temperature) +
              s(vegetation),
              data = dat,
              family = binomial())

p3 <- predict(object=preds,
             model=fm3,
             type = 'response')
plot(p2)
```

The second function can be used to develop a bias layer which can be included as probabilities of including background points in the generation of quasirandom background points.

\code{eip} (estimate inclusion probabilties)

@@

### References

Grafström, Anton, and Yves Tillé. "Doubly balanced spatial sampling with spreading and restitution of auxiliary totals." Environmetrics 24.2 (2013): 120-131.
